### 2019-04-19 Writing/response_to_reviewers.tex

\item{\textcolor{blue}{Along the lines set out by Reviewer 2, it would be wise to ensure that your metric is characterised as thoroughly as possible. In this regard, does it meet the following five conditions?}}

After working on the required proof of the metric, we figured out that the formerly proposed Character Difference had a non-expected behaviour of being non-linear.
We have now modified the metric to be linear (see illustration here: \url{https://github.com/TGuillerme/CharactersCorrelation/blob/master/Analysis/05-char.diffModification.pdf}).
We thus made some major changes to the manuscript that were not suggested by the reviewers below:
\begin{enumerate}
    \item We change the Character Difference metric to be a scaled Hamming Distance rather than the former proposed distance.
    \item We thus removed the supplementary material 1 wit the proof for the distance metric (since Hamming distances are well established distance metrics).
    \item We reran all the simulations from scratch with the inclusion of the updated Character Difference metric.
    \item For time constraint reason we reran 20 replicates (rather than 35 in the previous manuscript) for a total of 40 CPU years.
    \item We propose a new interpretation to the updated results.
\end{enumerate}

Of course, we also took all the reviewers comments on board (excluding the ones rendered obsolete by the modified metrics such as the proof for the four point condition).

We have also updated the Character Difference metric characterisation in the supplementary material 1.
However, since it is now simply a scaled Hamming distance, these demonstrations might be unwarranted for this revised version of the manuscript.

@@

\subsection{Major suggestions}

\begin{enumerate}

\item{\textcolor{blue}{1. To be a metric, there are several properties that need to be satisfied. The proposed function
is positively valued and symmetric (copied below from line 178): [...] But, the following metric property does not hold for all x; y:}}
\label{proof}

We have reviewed the formula of the CD metric following an unexpected non-linear behaviour (see comment \ref{metric_change}).
Furtheremore, we'd like to highlight the following points:

% We are really grateful that this reviewer picked up this flow in our mathematical demonstration of CD metric.
% This problem is due to a lack of indications from our side and we apologise for this.

### 2019-04-19 Writing/draft_highlighted.tex
changes to results part 1

The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology when maximising character difference leading to the smallest loss in phylogenetic information (median $NTS_{RF}$ = 0.956 and median $NTS_{Tr}$ = 0.839\hl{; see supplementary materials 3, Table 2}) followed by simply randomising the characters (median $NTS_{RF}$ = 0.762 and median $NTS_{Tr}$ = 0.628\hl{; see supplementary materials 3, Table 2}) and minimising the character difference (median $NTS_{RF}$ = 0.605 and median $NTS_{Tr}$ = 0.303\hl{; see supplementary materials 3, Table 2}).
There is a significant difference between all scenarios (maximising, minimising and randomising; \hl{\textit{p} = 0; Table} \ref{Tab_pooledscenarios_test}) with the highest probability of overlap being between maximising and randomising the character difference (Bhattacharrya Coefficient (BC) = 0.873 for the $NTS_{RF}$ and 0.908 for the $NTS_{Tr}$; Table \ref{Tab_pooledscenarios_test}) and the lowest probability between maximising and minimising the character difference (BC = 0.573 for the $NTS_{RF}$ and 0.614 for the $NTS_{Tr}$; Table \ref{Tab_pooledscenarios_test})

@@

\subsubsection{Number of characters}
This effect of the character difference is not dependent on the number of characters when looking at clade conservation (i.e. $NTS_{RF}$).
The median $NTS_{RF}$ was similar for 100, 350 and 1000 characters (0.730, 0.745, 0.767 respectively; see supplementary materials 3 \hl{Table 3}) with a significant difference only between 100 and 1000 and 350 and 1000 characters (\hl{\textit{p} = 0.001 and \textit{p} = 0.032 respectively}; Table \ref{Tab_pooledscharacters_test}).
The number of characters affects the character difference more in terms of taxon placement for a low number of characters (median $NTS_{Tr}$ for 100, 350 and 1000 characters equals 0.544, 0.693, 0.799 respectively - see supplementary materials 3 \hl{Table 3}) with a significant difference between 100 and 350 or 1000 characters (\hl{\textit{p} = 0, \textit{p} = 0 and \textit{p} = 0.014 respectively}; Table \ref{Tab_pooledscharacters_test}).
However, these differences have to be contrasted by a very high probability of overlap between each number of characters and metrics (Bhattacharrya Coefficient always $>0.95$) suggesting that the significant effects of the number of characters still leads to \hl{very} similar distributions.

>>>>

The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology when minimising character difference leading to the smallest loss in phylogenetic information (median $NTS_{RF}$ = 0.869 and median $NTS_{Tr}$ = 0.787\hl{; see supplementary materials 3, Table 3}) followed by randomising the characters (median $NTS_{RF}$ = 0.814 and median $NTS_{Tr}$ = 0.791\hl{; see supplementary materials 3, Table 2}) and maximising the character difference (median $NTS_{RF}$ = 0.698 and median $NTS_{Tr}$ = 0.754\hl{; see supplementary materials 3, Table 2}).
There is a significant difference between all scenarios \hl{in terms of clade conservation ($NTS_{RF}$)} (maximising, minimising and randomising; \hl{\textit{p} = 0; Table} \ref{Tab_pooledscenarios_test}) with the highest probability of overlap being between minimising and randomising the character difference (Bhattacharrya Coefficient (BC) = 0.954 for the $NTS_{RF}$; Table \ref{Tab_pooledscenarios_test}) and the lowest probability between maximising and minimising the character difference (BC = 0.848 for the $NTS_{RF}$; Table \ref{Tab_pooledscenarios_test}).
\hl{There is however no difference in terms of taxa displacement ($NTS_{Tr}$) between the three scenarios (with a BC > 0.98 for each comparison; Table \ref{Tab_pooledscenarios_test}).}

@@

\subsubsection{Number of characters}
\hl{This effect of the character difference affects clade conservation ($NTS_{RF}$) only by big differences in number of characters (100 \textit{vs.} 1000, median $NTS_{RF} = 0.727$ and $NTS_{RF} = 0.865$ respectively; Wilcoxon text $p = 0$ and $BC = 0.914$ \ref{Tab_pooledscharacters_test}).}
\hl{For closer differences of characters (i.e. 100 \textit{vs.} 350 and 350 \textit{vs.} 1000), Wilcoxon tests where different ($p = 0$) but with high distribution overlap ($BC > 0.96$).}

is not dependent on the number of characters when looking at clade conservation (i.e. $NTS_{RF}$).

The median $NTS_{RF}$ was similar for 100, 350 and 1000 characters (0.730, 0.745, 0.767 respectively; see supplementary materials 3 \hl{Table 3}) with a significant difference only between 100 and 1000 and 350 and 1000 characters (\hl{\textit{p} = 0.001 and \textit{p} = 0.032 respectively}; Table \ref{Tab_pooledscharacters_test}).
The number of characters affects the character difference more in terms of taxon placement for a low number of characters (median $NTS_{Tr}$ for 100, 350 and 1000 characters equals 0.544, 0.693, 0.799 respectively - see supplementary materials 3 \hl{Table 3}) with a significant difference between 100 and 350 or 1000 characters (\hl{\textit{p} = 0, \textit{p} = 0 and \textit{p} = 0.014 respectively}; Table \ref{Tab_pooledscharacters_test}).
However, these differences have to be contrasted by a very high probability of overlap between each number of characters and metrics (Bhattacharrya Coefficient always $>0.95$) suggesting that the significant effects of the number of characters still leads to \hl{very} similar distributions.
